Code Repository

### 2025-06-25 Code Repository.Rmd
Update Code Repository.Rmd
Wouldn't let me commit remotely

@@

```
Plotting theoretical null distribution of squared test statistic composite with eigenvalue weights
```{r}
ggplot(data = theoretical_null_squared, aes(x =chi_sq )) + geom_histogram(bins =250, colour = "lightsalmon", fill = "lightsalmon") + labs(x = "Squared Composite Test Statistic (QC)", y = "Frequency") +
  theme_minimal() + 
  geom_vline(xintercept = quantile(theoretical_null_squared$chi_sq, probs = c(0.95)), color = "red", linetype = "dashed", linewidth = 0.75) +
  annotate("text", x = quantile(theoretical_null_squared$chi_sq, 0.95), y = 25000, label = "95th Percentile", color = "red", angle = 90, vjust = -0.5, size = 5) +
  theme(panel.grid.minor = element_blank())+
  geom_vline( xintercept = study_squared_composite, color = "steelblue", size = 1 ) + 
  annotate("text", x = study_squared_composite, y = 25000, label = "Composite (QC = 29.2)", color = "steelblue", angle = 90, vjust = -0.5, size = 5)
```

@@

Make sure to remove ## for the code to run
```{r}
data <- tmpDat[,c("Label", "binary_group", "gw32", "gw38", "binary_insulin")]
inverted_data <- data[complete.cases(data), ]
inverted_data$binary_insulin <- ifelse(inverted_data$binary_insulin == 1, 0, 1)
inverted_data$gw32 <- inverted_data$gw32 * -1
inverted_data$gw38 <- inverted_data$gw38 * -1
inverted_data$baseline1 <- 0
inverted_data$baseline1[2] <-0.001
inverted_data$baseline2 <- 0
inverted_data$baseline2[4] <- 0.001

@@

## Power Estimation
Power estimation function
```{r}
power_function <- function(x, y, z, e, n) {
  summed_TS_power_result <- numeric(1000)  
  squared_TS_power_result <- numeric(1000)  
  EMERGE_composite_power <- numeric(1000)
  i <- 1
  repeat {
    placebo_group <- tmpDat[tmpDat$Group == "Placebo", ]
    placebo_group <- placebo_group[complete.cases(placebo_group[, c(x, y, z)]), ]
    sampled_placebo <- placebo_group[sample(nrow(placebo_group), n/2, replace = TRUE), ]

@@

    summed_TS_power_result[i] <- (t_test_result1$statistic + t_test_result2$statistic + t_test_result3$statistic)
    squared_TS_power_result[i]<-(((t_test_result1$statistic)^2)+((t_test_result2$statistic)^2)+((t_test_result3$statistic)^2))
    EMERGE_composite_power[i] <- t_test_result4$p.value
    i<- i +1
    if(i > 1000) break}
  return(as.numeric(c(sum(summed_TS_power_result<quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025)| summed_TS_power_result > quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975))/ 1000,## permutation test summed
         sum(squared_TS_power_result>quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95))/ 1000, ##permutation test squared
         sum(squared_TS_power_result > quantile(theoretical_null_squared$chi_sq, 0.95))/ 1000, ##asymptotic chi-square
         sum((EMERGE_composite_power< 0.05))/ 1000, ## EMERGE
         sum(summed_TS_power_result<quantile(null_normal_distribution, 0.025)| summed_TS_power_result > quantile(null_normal_distribution, 0.975))/ 1000)))
}
```

@@

>>>>

size<-100000
null_normal_distribution_small <- rnorm(size, mean = 0, sd = sqrt(variance))
write.csv(null_normal_distribution_small, "Theoretical Summed null distribution small.csv", row.names = F)
```
the p-value can be estimated using pnorm() for the asymptotic distribution of the composite (SC)
```{r}
p_value_pnorm <- 2*pnorm(study_summed_composite, mean =0, sd = sqrt(variance))
```

@@

size <- 100000
simulated_chi_sq_matrix_small <- matrix(rchisq(size * 3, df = 1), ncol = 3)
eigen_values_chi_sq_small<- simulated_chi_sq_matrix_small %*%eigen_values  
write.csv(eigen_values_chi_sq_small, "Theoretical Squared null distribution small.csv", row.names = F)
```
Plotting theoretical null distribution of squared test statistic composite with eigenvalue weights
```{r}
ggplot(data = theoretical_null_squared, aes(x =chi_sq )) + geom_histogram(bins =250, colour = "lightsalmon", fill = "lightsalmon") + labs(x = "Squared Composite Test Statistic (QC)", y = "Frequency") +
  theme_minimal() + 
  geom_vline(xintercept = quantile(theoretical_null_squared$chi_sq, probs = c(0.95)), color = "red", linetype = "dashed", linewidth = 0.75) +
  annotate("text", x = quantile(theoretical_null_squared$chi_sq, 0.95), y = 25000, label = "95th Percentile", color = "red", angle = 90, vjust = -0.5, size = 5) +
  theme(panel.grid.minor = element_blank())+
  geom_vline( xintercept = study_squared_composite, color = "steelblue", size = 1 ) + 
  annotate("text", x = study_squared_composite, y = 25000, label = "Composite (QC = 29.2)", color = "steelblue", angle = 90, vjust = -0.5, size = 5)
```

@@

Make sure to remove ## for the code to run
```{r}
data <- tmpDat[,c("Label", "binary_group", "gw32", "gw38", "binary_insulin")]
inverted_data <- data[complete.cases(data), ]
inverted_data$binary_insulin <- ifelse(inverted_data$binary_insulin == 1, 0, 1)
inverted_data$gw32 <- inverted_data$gw32 * -1
inverted_data$gw38 <- inverted_data$gw38 * -1

@@

## Power Estimation
Power estimation function
```{r}
power_function <- function(x, y, z, e, n) {
  summed_TS_power_result <- numeric(10000)  
  squared_TS_power_result <- numeric(10000)  
  EMERGE_composite_power <- numeric(10000)
  i <- 1
  repeat {
    placebo_group <- tmpDat[tmpDat$Group == "Placebo", ]
    placebo_group <- placebo_group[complete.cases(placebo_group[, c(x, y, z)]), ]
    sampled_placebo <- placebo_group[sample(nrow(placebo_group), n/2, replace = TRUE), ]

@@

    summed_TS_power_result[i] <- (t_test_result1$statistic + t_test_result2$statistic + t_test_result3$statistic)
    squared_TS_power_result[i]<-(((t_test_result1$statistic)^2)+((t_test_result2$statistic)^2)+((t_test_result3$statistic)^2))
    EMERGE_composite_power[i] <- t_test_result4$p.value
    i<- i +1
    if(i > 10000) break}
  return(as.numeric(c(sum(summed_TS_power_result<quantile(permuted_null_distributions.df$permutation_summed_composite, 0.025)| summed_TS_power_result > quantile(permuted_null_distributions.df$permutation_summed_composite, 0.975))/ 10000,## permutation test summed
         sum(squared_TS_power_result>quantile(permuted_null_distributions.df$permutation_squared_composite, 0.95))/ 10000, ##permutation test squared
         sum(squared_TS_power_result > quantile(theoretical_null_squared$chi_sq, 0.95))/ 10000, ##asymptotic chi-square
         sum((EMERGE_composite_power< 0.05))/ 10000, ## EMERGE
         sum(summed_TS_power_result<quantile(null_normal_distribution, 0.025)| summed_TS_power_result > quantile(null_normal_distribution, 0.975))/ 10000)))
}
```

@@

##Comparing Asymptotic and permutation null distributions
```{r}
asymptotic_null_summed_small <- read.csv("Theoretical Summed null distribution small.csv")
summed_composite_histogram.df <- data.frame(ts = c(permuted_null_distributions.df$permutation_summed_composite, asymptotic_null_summed_small$V1), group = rep(c("Permuted", "Theoretical"), c(length(permuted_null_distributions.df$permutation_summed_composite), nrow(asymptotic_null_summed_small))))

ggplot(data = summed_composite_histogram.df, aes(x =ts, fill = group )) + geom_histogram(bins =250, alpha = 0.5, position = "identity") + labs(x = "Summed Composite Test Statistic (SC)", y = "Frequency") +
  theme_minimal() + theme(panel.grid.minor = element_blank()) +
  scale_fill_manual(values = c("yellow", "skyblue"), name = "Method")
```

```{r}
asymptotic_null_sq_small <- read.csv("Theoretical Squared null distribution small.csv")

squared_composite_histogram.df <- data.frame(ts = c(permuted_null_distributions.df$permutation_squared_composite, asymptotic_null_sq_small$V1), group = rep(c("Permuted", "Theoretical"), c(length(permuted_null_distributions.df$permutation_squared_composite), nrow(asymptotic_null_sq_small))))

ggplot(data = squared_composite_histogram.df, aes(x =ts, fill = group )) + geom_histogram(bins =250, alpha = 0.5, position = "identity") + labs(x = "Squared Composite Test Statistic (QC)", y = "Frequency") +
  theme_minimal() + theme(panel.grid.minor = element_blank()) +
  scale_fill_manual(values = c("yellow", "skyblue"), name = "Method")
```
